---
format:
  html:
    embed-resources: true
---

## R and RStudio Assignment 1

In this assignment you will complete a variety of basic R/RStudio tasks with an emphasis on Tidyverse content. This assignment is drawn primarily from basic topics covered in our lectures and the **R for Data Science** textbook (link to the textbook in the course syllabus).

**Deliverable:** Each question in the assignment corresponds to a question in the Canvas quiz associated with this assignment.

```{r}
#| include: false
library(tidyverse) #our standard package for working in R
library(esquisse) #a visualization helper package (may not be needed)
library(nycflights13) #contains a dataset that we will need
```

------------------------------------------------------------------------

**Question 1**: The "nycflights13" package includes a dataset containing data on flights departing from three of New York City's airports (LaGuardia, JFK, and Newark). Use the line of code below to read in this dataset into a data frame called "diamonddata".

```{r}
flights = nycflights13::flights
```

Use an appropriate R command to answer the question: How many rows are in this dataset?

```{r}
nrow(flights)
```


**Question 2**: Use an appropriate R command to answer the question: How many columns are in the "flights" dataset?

```{r}
ncol(flights)
```


**Question 3**: Use the appropriate R function to select rows in the dataset with departure delays greater than 10 minutes (Hint: The variable that records departure delay is called "dep_delay").

```{r}
delay = flights %>% filter(dep_delay > 10)
```


How many flights in the dataset had a delay greater than 10 minutes?

82834

**Question 4**: Use the appropriate R function to select rows in the dataset with flights that departed from "JFK". (Hint: The variable that records departure airport is called "origin").

```{r}
kennedy = flights %>% filter(origin == "JFK")
```

How many flights in the dataset departed from JFK?
111279

**Question 5**: Create a new data frame with flights with "air_time" values greater than 180. Call this data frame "long".

```{r}
long <- flights %>% filter(air_time > 180)
```

How many rows are in this new data frame?

89476

**Question 6**: Using your new data frame from Question 5, determine the average "dep_delay" of flights in this data frame.

```{r}
mean(long$dep_delay)
```

Before proceeding, use the "read_csv" function to read-in the "InventoryData.csv" file into a data frame called "inventory". The CSV file is already uploaded to RStudio Cloud (you can see it in the Files pane at right).

```{r}
inventory <- read_csv("InventoryData.csv")
summary(inventory)
```


In the "inventory" data frame that you have just created, what type of variable is "Supplier"?

A. numeric\
B. integer\
C. factor\
D. character ++++

**Question 7** In the "inventory" data frame that you have just created, what is the standard deviation of the "On Hand" variable?

```{r}
sd(inventory$`On Hand`)
```


Hint: Be careful to write out On Hand inside backticks. A backtick is the character that looks like an apostrophe. It is typically on the same key as the \~ symbol. We have to do this because On Hand has a space in it (It's good practice to avoid spaces in variable names for this reason).

**Question 8**: What does the code below do?

```{r}
inventory = inventory %>% mutate(OnHandRatio = `On Hand` / `Annual Demand`)
```

A. Creates a new variable called "inventory"\
B. Creates a new variable called "OnHandRatio"\ +++
C. Creates a new data frame called "OnHandRatio"

**Question 10**: Using the "inventory" data frame, create a new data frame called "avg_cost" that contains the average "Cost per Unit (\$)" by each "Supplier" (let this quantity be in a variable called "SupplierAvgCost". Hint: Recall the summarize and group_by functions and the use of the backtick character. Your data frame should have only two columns:"Supplier" and "SupplierAvgCost".

What is the "SupplierAvgCost" of items from supplier A?

```{r}
avg_cost = inventory %>% group_by('Supplier') %>% summarise(SupplierAvgCost = mean('Cost per Unit ($)'))
```


------------------------------------------------------------------------
